\chapter*{Acknowledgements}
\label{cha:acknowledgements}
\markboth{Acknowledgements}{}
\addcontentsline{toc}{chapter}{Acknowledgements}

First, I would like to thank my supervisors, prof. Pierangelo Masarati and prof. Marco Morandini for giving me the opportunity to work on this fascinating topic and  letting me do it at my own pace. I had the chance to attend  to their courses and  that inspired me to improve my knowledge.
I would also thank them for their support and suggestions, even if these difficult times made everything a little more complicated. 

My special thanks go to the preCICE community, to Benjamin Uekermann, Gerasimos ``Makis'' Chourdakis and many others who were happy to exchange ideas and shared their knowledge and experience, helping me better understand the most difficult aspects of this subject.

Finally, but not least importantly, I owe my gratitude to Eleonora Colombo, who let me endlessly talk about my doubts and difficulties. She helped me put everything in perspective and gave me the motivation to pursue my goal, independently from everything else. Without her support this would not have been possible. 

%First and foremost, I would like to deeply thank my supervisor, Benjamin Uekermann, for
%his support. His vivid interest on the topic and his very helpful advices made my thesis a
%very pleasant learning experience.


%Additionally, I would like to thank Lucía Cheung Yau, whose excellent work I had the
%challenging mission to continue. Her aid in the beginning of my Thesis was invaluable. I
%also want to thank Babak Gholami for kindly providing us with the results of Lucía’s thesis,
%which was written in cooperation with SimScale GmbH.
%The tools I used in my thesis, mainly preCICE and OpenFOAM, are developed as
%free/open source software. They are also based on other free software projects, including
%the Doxygen documentation generator. Projects like these make me feel very glad of
%the free software community and I hope that my contribution will also be useful.
%The TUM English Writing Center helped me improve my understanding of the English
%language and the clarity of my prose. I am very thankful for the time they devoted to me
%during the Thesis Writers’ Workshop, as well as to our individual appointments.
%My Master’s studies were financially aided by a scholarship from the German Academic
%Exchange Service (DAAD) and I am very grateful of their support.
%Finally, I owe my deepest gratitude to my friends and family, whose endless support was
%crucial for my studies. Special thanks go to my parents and to my friend Myrto, who were
%there when their support was needed the most.