\chapter{Conclusions}
\label{cha:conclusions}
%\markboth{Conclusions}{}
%\addcontentsline{toc}{chapter}{Conclusions}

In this thesis, we  developed an adapter for MBDyn to enable fluid-structure interaction simulations with the coupling library preCICE.
The adaprer has been validated both qualitatively and quantitatively with several test cases.
This work gives insights into the basics of FSI simulations and allows to classify the implemented partitioned
coupling among the wide variety of coupling strategies

The adapter that we propose applies to a wider set of OpenFOAM solvers, is more
configurable, and can be plugged in at runtime without requiring any modifications to the
solver’s source code – an extra step that previously raised difficulties for the user.


Our proposed adapter not only allows for higher compatibility, but also makes the preparation
stage easier because it can simply be “plugged-in” at runtime. Up to now, this
required additional changes in the solver’s source code. This was achieved by wrapping the
adapter into a function object, a pattern also used by bundled post-processing tools.

Our adapter is also more flexible. The solver-specific parameters are read from the
OpenFOAM configuration files (dictionaries) and the user does not need to specify them
elsewhere. The boundary conditions on the interfaces are configured through the adapter’s
configuration file (YAML) and have been extended to support more solvers. The adapter
can now directly read any additionally required material properties from the case files,
instead of requiring modifications to the solver to read them. In case a solver uses different
names for the required fields, these are also configurable.

We validated the proposed adapter with multiple solvers, for two scenarios. The first
one was a flow over a heated plate, described in [19.]. In most of the cases, the result files
using the proposed adapter were identical to the files that were produced using the previous
adapters. In some cases, sporadic errors in the order of the results’ precision were observed.
In the case of the buoyantBoussinesqPimpleFoam, a problem was discovered and corrected
in the previous implementation, which was an example for the need for checkpointing in
implicit coupling. The second scenario was a shell-and-tube heat exchanger, a complex

This work makes preCICE compatible with a wide range of OpenFOAM solvers used
widely in both academia and industry. This brings preCICE not only closer to users but
also to developers. Being designed with extensibility in mind, our proposed adapter could
be equipped with additional boundary conditions for mechanical fluid-structure interaction
or other computational fields, making multi-physics simulations more accessible. In this
direction, the adapter could be ported to older versions and other variants of OpenFOAM.
Test cases should be developed, in order to make the continuous development of the adapter
for multiple OpenFOAM versions easier. Additionally, tutorials would help to reach new
users. In order to convince the OpenFOAM community, examples that highlight the advantages
of preCICE over the monolithic OpenFOAM solvers for conjugate heat transfer
should be presented. Developing additional modules for mechanical fluid-structure interaction
or acoustics would help to establish preCICE and OpenFOAM as standard components
of multi-physics simulations, leading to benefits for both communities




An adapter for linking the CFD solver SU2 with the multiphysics coupling tool preCICE is developed in
this work and validated both qualitatively and quantitatively with several testcases.
This work gives insights into the basics of FSI simulations and allows to classify the implemented partitioned
coupling among the wide variety of solver strategies. The integration of the adapter into the code
structure of SU2 is explained, as well as its implementation itself. The coupling approach is validated
via generic two- and three-dimensional testcases, and the well-known numerical benchmark scenario FSI3
([39]). The adapter is capable of handling three-dimensional, real-world examples. This is exemplified
by simulating a long, slender cylinder under turbulent flow conditions. The practical background of this
simulation is research on brush seals for e.g. gas turbines. However, time did not suffice to obtain meaningful
simulation results in the scope of this thesis. Thus, work on this topic will be continued. Once
feasible single-cylinder simulation results are achieved, scenarios involving multiple cylinders in a channel
should be simulated in order to include possible mutual interactions of these structures. Thereby, the
real conditions in brush seals can be approximated more precisely. In its current version, the adapter
is limited to a single wet surface in the SU2 mesh file. While it is technically possible to define the
surface of multiple cylinders as a single wet surface, it is more accurate to extend the adapter such that
an arbitrary number of separate FSI interfaces can be defined. Moreover, it might be useful to integrate
a load ramping functionality in the adapter, such that at the beginning of an FSI simulation the forces
exerted on the structure are reduced by a factor, which is consequently increased to recover the real
force values after a certain number of time steps. This might help to stabilize simulations as the initially
occurring, large displacements can be reduced by this method. Recently, several RBF mapping methods
have been included in preCICE, which can be used for parallel simulations. Applying these elaborate
interpolation procedures instead of NN mapping allows to use finer solid meshes without encountering
oscillatory forces at the wet surface, such that simulation results are expected to significantly improve
(compared to simulations done with NN mapping). Furthermore, the spatial oscillations at the FSI interface
are expected to vanish. Consequently, further testing with the FSI3 benchmark scenario will be
done in order to verify this.
The adapter is implemented such that it perfectly integrates into SU2, allowing to reuse characteristic
functionalities like turbulence modeling, convergence acceleration via dual time stepping and multi-grid
functionalities. Moreover, usage of preCICE for coupled FSI simulations can be configured via the native
SU2 configuration file. A single executable of SU2 can be used for different FSI scenarios and even
fluid-only computations without the need for recompilation. The adapter allows for inter- and intrafield
parallel simulations. While no ALE implementation is currently available for the incompressible solver
in SU2, the adapter is already prepared for this feature, possibly extending the FSI capabilities to the
incompressible regime in the future.
Although intrinsic FSI functionalities were recently added to SU2, the coupling with preCICE offers more
flexibility, as even commercial partner solvers can be chosen for multiphysics simulations. Moreover, the
coupling algorithms and data mapping methods of preCICE are more elaborate. The possibility of
running fully-parallel simulations outperforms the intrinsic FSI capabilities with regard to HPC.
Another practically very useful outcome of this work is the description of the procedure for integrating the
adapter into SU2 and building it with preCICE. Thus, users can easily extend SU2 to FSI via preCICE
guided by this thesis, which addresses both SU2 and preCICE communities. The adapter will soon be
included in the preCICE repository.